---
title: "Dosage data and analyses brief"
# subtitle: ""
description: "This document shows the analyses of the dosage paper."

author: 
  name: 'Emiel Schoneveld'
  affiliations:
    name: 'University of Amsterdam'

keywords: "Dosage, Time, Words, Descriptives, Analyses"
date: today

execute: 
  warning: false
  message: false
  echo: false

project:
     execute-dir: project

editor: source
number-sections: true
format:
  html:
    embed-resources: true # used for making a standalone file, slow render but perfect for sharing
    toc: true
    code-fold: false
    # code-summary: "Show the code"
    title-block-banner: "#bc0031" # title block background color
    title-block-banner-color: "#fffff" # title block font color
    backgroundcolor: "#F5F5F3"
    fontcolor: "#1B1918"
---

```{r echo=FALSE, include=FALSE}
# Preperation syntax
## Load libraries
library(here)
library(tidyverse)
library(vtable)
library(DT)
library(reshape2)
library(plotly)
library(scatterPlotMatrix)
library(shiny)

## source descriptive syntax and analyses draft
source('08_analysis_mediation.R')
source('09_analysis_interaction.R')
```

# Mediation analysis
## Model description
This model is used to investigate whether the wordreading improvements are caused by the number of words students read and/or the amount of time they have practiced. Words exposures is defined ass the number of times a students was presented with a word.
The model displayed below was fitted with and without the direct effect of time on post.

::: callout-note
The covariances are omitted in the figure for legibility.
:::

```{r echo=FALSE, include=TRUE}
# Plot model
lavaanPlot(
  fit_partialmediation_discrete,
  graph_options = list(rankdir = "TB")
)
```

## Results
The models were fitted for five outcome measures. Below the standardized estimates of the regression parameters and the full vs partial mediation anova test.

::: {.callout-note}
Greyed out boxes are omitted regression paths because of full mediation
:::


```{r echo=FALSE, include=TRUE}
p_mediation_formatted
```

# Interaction analysis
## Model description
Using an OLM linear model the following model was fitted to check the influence of accuracy. The effect of time on post was omitted for discrete and serial because it was fully mediated by words_exposures.

```{r echo=FALSE, include=TRUE}
# specify model
mod_lm_visualisation <- 'fluency_post ~ fluency_pre + time + words_exposures + accuracy_anytry + exposuresXaccuracy'

# fit dummy model
fit_lm_visualisation <- sem(
  mod_lm_visualisation,
  data |> 
    mutate(
      fluency_post = fluency_DMT_post,
      fluency_pre = fluency_DMT_pre,
      exposuresXaccuracy = accuracy_anytry*words_exposures,
    )
)

# Plot
lavaanPlot(
  fit_lm_visualisation
)


```

## The results
The models were fitted for three outcome measures. Below the standardized regression estimates.

::: {.callout-note}
Greyed out boxes are omitted regression paths because of full mediation in the previous analysis.
:::

```{r echo=FALSE, include=TRUE}
p_interaction_formatted
```