---
title: "Dosage data and analyses brief"
# subtitle: ""
description: "This document shows the descriptive statistics and potential analyses of the dosage paper."

author: 
  name: 'Emiel Schoneveld'
  affiliations:
    name: 'University of Amsterdam'

keywords: "Dosage, Time, Words, Descriptives, Potential Analyses"
date: today

execute: 
  warning: false
  message: false
  echo: false

project:
     execute-dir: project

editor: source
number-sections: true
format:
  html:
    embed-resources: true # used for making a standalone file, slow render but perfect for sharing
    toc: true
    code-fold: false
    # code-summary: "Show the code"
    title-block-banner: "#bc0031" # title block background color
    title-block-banner-color: "#fffff" # title block font color
    backgroundcolor: "#F5F5F3"
    fontcolor: "#1B1918"
---

```{r echo=FALSE, include=FALSE}
# Preperation syntax
## Load libraries
library(here)
library(tidyverse)
library(vtable)
library(DT)
library(reshape2)
library(plotly)

## source descriptive syntax and analyses draft
source('07_descriptive_analyses.R')
source('08_analyses_drafts.R')

## Remove mod objects
rm(list = ls()[str_detect(ls(), 'mod')])
```

# Summary statistics
## Table
This table shows the summary statistics for all practice variables and DMT scores.
```{r echo=FALSE, include=TRUE}
# Make variable description
var_description <- tribble(
  ~Variable, ~Description,
  'frequency', 'average sessions per week',
  'duration', 'number of weeks practiced',
  'length', 'average session length in minutes, defined as time between start of the first exercise and ending of the last exercise of the day',
  'cii_sessions', 'number of sessions a student completed',
  'cii_time', 'total hours of practice defined as the sum of all session length',
  'cii_readingtime', 'total hours of reading time, defined as the sum of all exercise duration, this excludes breaks inbetween exercises',
  
  'cii_words_exposures', 'number of times the student was exposed to any word',
  'cii_words_unique', 'number of times the student was exposed to a new word, contain duplicates when word were presented in multiple lessons',
  'cii_words_accurate', 'number of times the student read a word accurate at their first attempt, if the tutor played the audiorecording of the word the reading was regarded inaccurate',
  'cii_words_audioplays', 'number of times the audiorecording of words were played',
  'cii_words_attempts', 'number of times the student read words a second or a third time after an incorrect reading',
  'accuracy', 'percentage of words read accurately at the first attempt, defined as cii_words_accurate/cii_words_unique',
) 

# Print variable description
var_description |> kable()
```


```{r echo=FALSE, include=TRUE}
# Make summary tables
table_fullsample <- data |> 
  dplyr::select(
    contains('practice'),
    contains('wordreading')
  ) |> 
  dplyr::select(
    !(contains('missing') | contains('outlier') | contains('invalid'))
  ) |> 
  sumtable(
    out = "return"
  ) |> as_tibble() |> 
  mutate(
    grade = 'all', .before = 1
  )

table_grade2 <- data |> 
  filter(
    grade == "grade_2"
  ) |> 
  dplyr::select(
    contains('practice'),
    contains('wordreading')
  ) |> 
    dplyr::select(
    !(contains('missing') | contains('outlier') | contains('invalid'))
  ) |> 
  sumtable(
    out = "return",
  ) |> 
  as_tibble() |> 
  mutate(
    grade = 'grade_2', .before = 1
  )

table_grade3 <- data |> 
  filter(
    grade == "grade_3"
  ) |> 
  dplyr::select(
    contains('practice'),
    contains('wordreading')
  ) |> 
  dplyr::select(
    !(contains('missing') | contains('outlier') | contains('invalid'))
  ) |> 
  sumtable(
    out = "return",
  ) |> as_tibble() |> 
  mutate(
    grade = 'grade_3', .before = 1
  )

table_grade4 <- data |> 
  filter(
    grade == "grade_4"
  ) |> 
  dplyr::select(
    contains('practice'),
    contains('wordreading')
  ) |> 
  dplyr::select(
    !(contains('missing') | contains('outlier') | contains('invalid'))
  ) |> 
  sumtable(
    out = "return",
  ) |> as_tibble() |> 
  mutate(
    grade = 'grade_4', .before = 1
  )

# Combine summary tables
table_all <- bind_rows(
  table_fullsample,
  table_grade2,
  table_grade3,
  table_grade4
) |> 
  mutate(
    grade = as_factor(grade),
    Variable = str_remove(Variable, 'practice_'),
    Variable = as_factor(Variable)
  )

# Print table as interactive table
table_all |> 
  datatable(
    filter = "top"
  )
```

## Histograms
These histograms show the distributions of the practice variables.

```{r echo=FALSE, include=TRUE}
# Plot histograms
p_histos <- data |> 
  dplyr::select(
    contains('cii_'),
    practice_accuracy,
    contains('score') & !contains('mid')
  ) |> 
  pivot_longer(
    contains('e'),
    names_to = "variable",
    values_to = "value"
  ) |> 
  mutate(
    variable = str_remove(variable, 'practice_') |> as_factor()
  ) |> 
  ggplot(
    aes(
      x = value,
    )
  ) +
  geom_histogram() +
  facet_wrap(
    ~variable,
    scales = 'free',
    ) +
  theme(
    panel.spacing.y = unit(0.5,"lines")
  )

# plot
p_histos |> print()

```

# Correlations
This plot shows the correlations between the practice variables and wordreading scores.

```{r echo=FALSE, include=TRUE}
# Show correlation
p_cors <- cors |> 
  round(1) |> 
  melt() |> 
  mutate(
    Var1 = str_remove(Var1, 'practice_') |> as_factor(),
    Var2 = str_remove(Var2, 'practice_') |> as_factor(),
  ) |> 
  drop_na(
    value
  ) |> 
  ggplot(
    aes(
      x = Var1,
      y = Var2,
      fill = value
    )
  ) +
  geom_tile(color = "white") +
  geom_text(
    aes(label = value),
    color = "black",
    size = 4
  ) +
  scale_fill_gradient2(
    low = "red",
    high = "darkgreen", 
    mid = "white",
    midpoint = 0,
    limit = c(-1,1),
    space = "Lab",
    name="Correlation"
    ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 30,
      vjust = 1,
      hjust = 1
    )
  )

ggplotly(p_cors)
```

# Analyses
For the analyses we have three possible models.

## Model 1: Mediation model
### Model description
This model is used to investigate whether the wordreading improvements are caused by the number of words students read and/or the amount of time they have practiced. Words unique is defined ass the number of times a students was presented with a new word.
The model displayed below shows that time predicts words which in turn predicts post reading score. In Grade 2 and Grade 4, time also has an additional negative effect on post reading score. In Grade 3, time does not have any impact on post reading after controlling for the amount of words students have read.
When the total effect is calculated for time on post it shows only a significant positive effect for Grade 2.

```{r echo=FALSE, include=TRUE}
# Plot model
lavaanPlot(
  fit_partialmediation,
  labels = list(
    practice_cii_time = "Time",
    practice_cii_words_unique = 'Words unique',
    wordreading_score_pre = 'Fluency pre',
    wordreading_score_post = 'Fluency post'
  ),
  coefs = T,
  # covs = T,
  digits = 1,
  stars = 'regress',
  graph_options = list(
    rankdir = "LR"
  )
)
```
::: callout-note
The covariances are omitted in the figure for legibility.
:::

### Mediation analysis
Omitting the direct effect of time on post reading leads to significant decrease in model fit.
```{r echo=FALSE, include=TRUE}
# Print mediation comparison
anova(fit_partialmediation, fit_fullmediation)
```

## Model 2: Accuracy 
### Model description
This model is used to investigate how accuracy during training impacts students' improvements after training. Accuracy is defined as the percentage of times a students read a flashed word correctly at their first attempt without the help of an audioplay.

The model displayed below shows that only in Grade 2 accuracy plays a positive role. Students who are more accurate during training improve their word reading fluency more. However, there is no significant interaction effect between words and accuracy. In summary, for an intervention in which students are predominantly very accurate (above 80%), all remaining differences in accuracy seem relatively unimportant.

```{r echo=FALSE, include=TRUE}
# Plot model
lavaanPlot(
  fit_wordstimeaccuracy_partialmediation,
  labels = list(
    practice_cii_time = "Time",
    practice_cii_words_unique = 'Words unique',
    practice_accuracy = 'Accuracy',
    wordreading_score_pre = 'Fluency pre',
    wordreading_score_post = 'Fluency post'
  ),
  coefs = T,
  # covs = T,
  digits = 1,
  stars = 'regress',
  graph_options = list(
    rankdir = "LR"
  )
)
```
::: callout-note
The productterm is allowed to covary with all variables to fit a saturated model. The covariances are omitted in the figure for legibility.
:::

### Mediation analysis
Omitting the direct effect of time on post reading does not lead to significant decrease in model fit. In this model, the effect of time of post reading fluency is completely mediated through words unique and accuracy. However, the resulting model is of inadequate fit.
```{r echo=FALSE, include=TRUE}
# Print mediation comparison
anova(fit_wordstimeaccuracy_partialmediation, fit_wordstimeaccuracy_fullmediation)
print('Fit of full mediation model')
fitmeasures(fit_wordstimeaccuracy_fullmediation, 
            c('df', 'chisq', 'pvalue', 'cfi', 'rmsea'))
```

## Model 3: Attempts model
### Model description
This model is used to investigate accurate readings and attempted readings evoke fluency improvements. Words accurate is defined as the number of words that the students has read accurately at their first attempt. Words attempts is defined as the number of attempts a student has performed at reading a word that was not a first and correct reading. These two variables are not defined by each other.

The model displayed below shows that only in Grade 2 the attempts at reading after an inaccurate reading impact fluency improvements. Grade 2 students who have attempted more readings after a first incorrect reading show smaller improvements.

```{r echo=FALSE, include=TRUE}
# Plot model
lavaanPlot(
  fit_accuratewordstime_partialmediation,
  labels = list(
    practice_cii_time = "Time",
    practice_cii_words_accurate  = 'Words accurate',
    practice_cii_words_attempts  = 'Words attempts',
    wordreading_score_pre = 'Fluency pre',
    wordreading_score_post = 'Fluency post'
  ),
  coefs = T,
  # covs = T,
  digits = 1,
  stars = 'regress',
  graph_options = list(
    rankdir = "LR"
  )
)
```

::: callout-note
The covariances are omitted in the figure for legibility.
:::

### Mediation analysis
Omitting the direct effect of time on post reading leads to significant decrease in model fit.
```{r echo=FALSE, include=TRUE}
# Print mediation comparison
anova(fit_accuratewordstime_partialmediation, fit_accuratewordstime_fullmediation)
```