---
title: "Make time count! Exposure to words, not hours spent practicing, drives reading progress"
subtitle: "A data report"
description: "This report provides a detailed overview of the data used for the SSSR poster. In addition it provides the estimates derived from both analyses."

author: 
  name: Emiel Schoneveld
  email: r.e.schoneveld@uva.nl
  orcid: 0000-0001-5679-6616
  affiliations:
    name: 'University of Amsterdam'

keywords: "Intervention, Dosage, Time, Words, Accuracy"
date: today

execute: 
  warning: false
  message: false
  echo: false

project:
     execute-dir: project

editor: source
number-sections: true
format:
  html:
    embed-resources: true # used for making a standalone file, slow render but perfect for sharing
    toc: true
    code-fold: false
    # code-summary: "Show the code"
    title-block-banner: "#bc0031" # title block background color
    title-block-banner-color: "#fffff" # title block font color
    backgroundcolor: "#F5F5F3"
    fontcolor: "#1B1918"
---

```{r echo=FALSE, include=FALSE}
# Preperation syntax
## Load libraries
library(here)
library(tidyverse)
library(scatterPlotMatrix)
library(DT)

## source descriptive syntax and analyses draft
# source('07_descriptive_analyses.R')
source('08_analysis_mediation.R')
source('09_analysis_interaction.R')

## Select variables and rename
data_wide <- data |> 
  dplyr::select(
    grade, time, words_exposures, accuracy_anytry, contains('fluency') & !(contains('LED') | contains('pseudo'))
  ) |> 
  rename(
    'words' = words_exposures,
    'accuracy' = accuracy_anytry
  ) |> 
  mutate(
    grade = case_when(
      grade == 'grade_2' ~ 'Grade 2',
      grade == 'grade_3' ~ 'Grade 3',
      grade == 'grade_4' ~ 'Grade 4',
    ),
    accuracy = accuracy*100 |> round(2)
  ) |> 
  rename_with(~ str_remove(.x, 'fluency_')) |> 
  rename_with(~ str_replace(.x, '_', '-')) |> 
  rename_with(~ str_replace(.x, 'serial', 'serial_exp')) |> 
  rename_with(~ str_replace(.x, 'discrete', 'discrete_exp')) |> 
  rename_with(~ str_replace(.x, 'DMT', 'serial_normed'))

## Pivot data
data_long <- data_wide |> 
  pivot_longer(
    contains('-'),
    names_to = c('task', 'measurement'),
    names_sep = '-',
    values_to = 'fluency'
  )

## Make factors
data_long <- data_long |> 
  mutate(
    task = factor(task, levels = c('serial_normed', 'serial_exp', 'discrete_exp')),
    measurement = factor(measurement, levels = c('pre', 'post')),
  )
```

# Descriptive statistics
## Descriptive table
```{r}
# Make summary table per grade
summary_table_grouped <- data_wide |> 
  pivot_longer(
    -grade,
    names_to = 'outcome',
    values_to = 'value'
  ) |> 
  group_by(
    grade, outcome
  ) |> 
  summarise(
    N = n(),
    N_missing = sum(is.na(value)),
    Min = min(value, na.rm = T) |> round(2),
    Max = max(value, na.rm = T) |> round(2),
    Mean = mean(value, na.rm = T) |> round(2),
    SD = sd(value, na.rm = T) |> round(2),
  )

# Make summary table for full sample
summary_table_fullsample <- data_wide |> 
  pivot_longer(
    -grade,
    names_to = 'outcome',
    values_to = 'value'
  ) |> 
  group_by(
    outcome
  ) |> 
  summarise(
    N = n(),
    N_missing = sum(is.na(value)),
    Min = min(value, na.rm = T) |> round(2),
    Max = max(value, na.rm = T) |> round(2),
    Mean = mean(value, na.rm = T) |> round(2),
    SD = sd(value, na.rm = T) |> round(2),
  ) |> 
  mutate(
    grade = 'All grades'
  )

# Bind summary tables
summary_table <- bind_rows(
  summary_table_fullsample,
  summary_table_grouped
) |> 
  mutate(
    grade = factor(grade, levels = c('All grades', 'Grade 2', 'Grade 3', 'Grade 4')),
    measurement = str_extract(outcome, "(?<=-).*"),
    outcome = if_else(
      str_detect(outcome, '-'),
      str_extract(outcome, ".*(?=-)"),
      outcome
    )
  ) |> 
  relocate(
    outcome, grade, measurement, N, N_missing, Min, Max, Mean, SD
  ) |> 
  arrange(
    outcome,
    grade
  )

# Render summary table interactively
datatable(
  summary_table
)
```

## Histograms and scatterplots
### Pre & post fluency
```{r}
#| column: screen
#| 
data_wide |> 
  dplyr::select(
    grade, contains('post')
  ) |> 
  scatterPlotMatrix(
    categorical = list(grade = sort(unique(data_wide$grade))),
    zAxisDim = 'grade',
    rotateTitle = T,
    width = 1400,
    height = 1400
  )

```

### Practice measures
```{r}
#| column: screen
#| 
data_wide |> 
  dplyr::select(
    time, words, accuracy, grade
  ) |> 
  scatterPlotMatrix(
    categorical = list(grade = sort(unique(data_wide$grade))),
    zAxisDim = 'grade',
    rotateTitle = T,
    width = 1400,
    height = 1400
  )
```

### Post & time
```{r}
#| column: screen
#| 
data_wide |> 
  dplyr::select(
    time, grade, contains('post')
  ) |> 
  scatterPlotMatrix(
    categorical = list(grade = sort(unique(data_wide$grade))),
    zAxisDim = 'grade',
    rotateTitle = T,
    width = 1400,
    height = 1400
  )

```

### Post & words
```{r}
#| column: screen
#| 
data_wide |> 
  dplyr::select(
    words, grade, contains('post')
  ) |> 
  scatterPlotMatrix(
    categorical = list(grade = sort(unique(data_wide$grade))),
    zAxisDim = 'grade',
    rotateTitle = T,
    width = 1400,
    height = 1400
  )

```

### Post & accuracy
```{r}
#| column: screen
#| 
data_wide |> 
  dplyr::select(
    accuracy, grade, contains('post')
  ) |> 
  scatterPlotMatrix(
    categorical = list(grade = sort(unique(data_wide$grade))),
    zAxisDim = 'grade',
    rotateTitle = T,
    width = 1400,
    height = 1400
  )

```

# Mediation analysis
## Model description
This model is used to investigate whether the reading fluency improvements are caused by the number of words students read and/or the amount of time they have practiced.
The model displayed below was fitted with and without the direct effect of time on post to test for full or partial mediation of the time effect on post fluency via words.

::: callout-note
The covariances are omitted in the figure for legibility.
:::

```{r echo=FALSE, include=TRUE}
# Plot model
lavaanPlot(
  fit_partialmediation_discrete,
  graph_options = list(rankdir = "TB")
)
```

## Results
The models were fitted for three outcome measures. Below the standardized estimates of the regression parameters and the full vs partial mediation anova test.

::: {.callout-note}
Time -> post is empty because the effect of time on post was fully mediated by words in all three models.
:::


```{r echo=FALSE, include=TRUE}
#| column: screen
#| fig-width: 12

p_mediation_formatted

```

# Interaction analysis
## Model description
Using an OLM linear model the following model was fitted to check the influence of accuracy. The effect of time on post was omitted because it was fully mediated by words.

```{r echo=FALSE, include=TRUE}
# specify model
mod_lm_visualisation <- 'fluency_post ~ fluency_pre + time + words_exposures + accuracy_anytry + exposuresXaccuracy'

# fit dummy model
fit_lm_visualisation <- sem(
  mod_lm_visualisation,
  data |> 
    mutate(
      fluency_post = fluency_DMT_post,
      fluency_pre = fluency_DMT_pre,
      exposuresXaccuracy = accuracy_anytry*words_exposures,
    )
)

# Plot
lavaanPlot(
  fit_lm_visualisation
)
```

## The results
The models were fitted for three outcome measures. Below the standardized regression estimates.

```{r echo=FALSE, include=TRUE}
#| column: screen
#| fig-width: 12
p_interaction_formatted

```
